Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-02-21 appendix.tex
I add section of step-by-step linearizability. Now Appendix B is the proof of step-by-step linearizability. In this process, proofs of two lemmas are changed.

@@

@@

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.

@@

@@

@@

According to $\textit{MS}(\textit{PQ}_1^{>})$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than priority of any method event in $u \cdot v \cdot w$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$. Let $S_1$ be the set of method events of operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all method events in $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$. It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

@@

\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.

@@

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed

\end {proof}

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] if $e$ is linearizable w.r.t. $\textit{MS}_{\textit{PQ}_1^{>}}$ ($R \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>},\textit{PQ}_2^{=} \}$) with witness $x$, we have: $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.

\item[-] if $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, we have:

$e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{itemize}

@@

@@

\section{Step-by-Step Linearizability of Priority Queues}
\label{sec:step-by-step linearizability of priority queues}

\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
\label{lemma:PQ as multi in MRpri for history}
Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability. $h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
\end{restatable}

The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.

To prove the $\textit{if}$ direction. Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$. Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$. By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$. Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.

The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$. $h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.

\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$. $h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.

\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$. $h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$. $h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$. $h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
\end{itemize}

Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$. Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$. It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$. Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$. This completes the proof of the $\textit{if}$ direction. \qed
\end {proof}

To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.

\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
\label{lemma:Rpr2 is step-by-step linearizability}

$R_{\textit{pr2}}$ is step-by-step linearizability.

\end{restatable}

Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$. Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.

According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$. Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$. Let $\textit{OP}'_x$ be the set of operations in $l_t$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

>>>>

\subsection{Proof of Lemma \ref{lemma:PQ is closed under projection}}

@@

\subsection{Proof of Lemma \ref{lemma:PQ as multi in MRpri for sequence}}

@@

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $e$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing one unmatched $\textit{put}$ which use the item last putted in $S$.

@@

\subsection{Proof of Lemma \ref{lemma:PQueue is step-by-step linearizability}}

@@

Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$) between each pair of $\textit{cal}(\textit{put},a)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$). The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.

@@

According to $\textit{MS}(\textit{PQ}_1^{>})$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than priority of any method event in $u \cdot v \cdot w$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$. Let $S_1$ be the set of operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all method events of operations in $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$. It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

@@

\item[-] If $o_1 = \textit{rm}(x) \wedge (o_2 \in  l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$, or $o_1 \in L'_R \wedge ( o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x) )$, then $l$ contradicts with happen before relation of $h$.

@@

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{PQ1EqualisStepByStepLinearizability}
\label{lemma:PQ1Equal is step-by-step linearizability}
If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_1^{=})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{restatable}

Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{=}$ with witness $x$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.

Since $h' \sqsubseteq l'$, there exists a sequence $e_{l'}$ generated from $e$ by adding method events of $e \setminus x$, and $e_{l'}$ satisfies that (1) the projection of $e_{l'}$ into call and return action is $e$, (2) the projection of $e_{l'}$ into method events is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $e$ (if their item is not $x$), there exists a method event $o$, such that $c$ before $o$ in $h'_{l'}$, $o$ before $r$ in $h'_{l'}$, and $o$ is the method event of $c$ and $r$. Intuitively, $e_{l'}$ is obtained from $h$ by adding linearization points according to $l'$.

According to $\textit{PQ}_1^{=}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ and $\textit{rm}$ in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ with priority $\textit{pri}_x$ is in $u$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'$, such that the linearization point of every method event in  $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$. Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into method events of $E_{px}$ that are not in $l'_{\textit{L0}}$. Let $E_{l_t}$ be the set of method events in $l_t$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into method events of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

As in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$. It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$. Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$. Since all $\textit{put}$ operations with priority $\textit{pri}_x$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$. Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' \in \textit{PQueue}$.

To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in l_t)$, then $l'$ contradicts with happen before relation of $h$.

    Else, it is the case that $o_1 \in l_t \wedge o_2 \in l'_{L0}$. We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$. Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.

    Else, it is the case that $o_2 \in l_t$. Since $o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e$. Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{PQ2LarisStepByStepLinearizability}
\label{lemma:PQ2Lar is step-by-step linearizability}
If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_2^{>})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{restatable}

According to $\textit{PQ}_2^{>}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$), and all other $\textit{put}$ with priority $\textit{pri}_x$ is in $u$. Let $S_1$ be the set of operations of $e$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all method events of operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$. Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.

Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$. It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{PQ2EqualisStepByStepLinearizability}
\label{lemma:PQ2Equal is step-by-step linearizability}
If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_2^{=})$ with witness $x$, then $e \setminus x \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, and $l$ be an sequential execution such that $h \sqsubseteq l$ and $l$ matches $\textit{PQ}_1^{>}$ with witness $x$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.

According to $\textit{PQ}_2^{=}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ in $u \cdot v$, (4) all matched $\textit{put}$ with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.

Similarly as in Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, we can obtain a sequence $e_{l'}$ from $e$ by adding linearization points according to $l'$. Let $l'_1$ be the shortest prefix of $l'$, such that the linearization point of every method event in $l'_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$. Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$. Let $l'_2$ be the projection of $l'$ into method events of $E_{px}$ that are not in $l'_1$. Let $l_x' = l'_1 \cdot l'_2 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_3$, where $l'_3$ is the projection of $l'$ into method events that are not in $l'_1 \cdot l'_2$.

Let $l_x$ be generated from $l'_x$ by erasing $\textit{put}(x,\textit{pri}_x)$. Since (1) $l_x$ can be obtained from $l'$ by making some $\textit{put}$ happen earlier, (2) for each $\textit{rm}(a)$ in $l'$, its corresponding $\textit{put}(a,\_)$ must before it in $l'$, and (3) there are no $\textit{rm}(\textit{empty})$ in $l'$, we can see that $l_x \in \textit{PQueue}$. Since $\textit{pri}_x$ is the maximal priority in $l$ and there are no $\textit{rm}(\textit{empty})$ in $l$, we can see that $l'_x \in \textit{PQueue}$.

Let us prove that $h \sqsubseteq l'_x$ by contradiction. Assume that there exists operations $o_1,o_2$, such that $o_1$ happens before $o_2$ in $h$, and $o_2$ is before $o_1$ in $l'_x$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1, o_2 \in l'_1$, or $o_1, o_2 \in l'_2$, or $o_1, o_2 \in l'_3$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_2 \in l'_1 \wedge o_1 \in l'_2$: Then by construction of $l'_1$, it is easy to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.

\item[-] $o_2 \in l'_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$: Then by construction of $l'_1$, it is easy to see that $o_2 \notin l'_1$, which contradicts that $o_2 \in l'_2$.

\item[-] $o_2 \in l'_1 \wedge o_1 \in l'_3$: Then by construction of $l'_1$, it is easy to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.

\item[-] $o_2 \in l'_2 \wedge o_1 = \textit{pt}(x,\textit{pri}_x)$: It is easy to see that each method event in $l'_2$ is a $\textit{put}$ with priority $\textit{pri}$. By $l'$ we know that $\textit{pt}(x,\textit{pri}_x)$ does not happen before any of them.

\item[-] $o_2 \in l'_2 \wedge o_1 \in l'_3$: It is easy to see that each method event in $l'_2$ is a $\textit{put}$ $o$ with priority $\textit{pri}$ and $o$ interleaves with $\textit{put}(x,\textit{pri}_x)$. Since $o_1$ happens before $o_2$, it is obviou that the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.

\item[-] $o_2 = \textit{put}(x,\textit{pri}_x) \wedge o_1 \in l'_3$: It is not hard to see that $o_1 \in l'_1$, which contradicts that $o_1 \in l'_3$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq l'_x$. \qed
\end {proof}

\begin{restatable}{lemma}{PQ3isStepByStepLinearizability}
\label{lemma:PQ3 is step-by-step linearizability}
If concurrent execution $e$ is linearizable w.r.t. $\textit{MS}(\textit{PQ}_3)$ and $o$ is a $\textit{rm}(\textit{empty})$ event, then $e \setminus o \sqsubseteq \textit{PQueue} \Rightarrow e \sqsubseteq \textit{PQueue}$.
\end{restatable}

\begin {proof}
Let $h$ be the data-differentiated history of $e$, $l$ be an sequential execution such that $h \sqsubseteq l$, $l$ matches $\textit{PQ}_3$ and $o$ is a $\textit{rm}$ method event in $h$. Let $h'=h \setminus o$ and assume that $h' \sqsubseteq l' \in \textit{PQueue}$.

According to $\textit{PQ}_3$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{rm}(\textit{empty}) \cdot v$, where the $\textit{put}$ operations and $\textit{rm}$ in $u$ are matched.

Let $E_L$ be the set of method events in $u$ and $E_R$ be the set of method events in $v$. Let $l'_L = l' \vert_{E_L}$ and $l'_R = l' \vert_{E_R}$. Let sequence $l'' = l'_L \cdot o \cdot L'_R$. Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u$ contains matched $\textit{put}$ and $\textit{rm}$, we know that $l'_L \in \textit{PQueue}$ and the the priority queue is empty after executing $l'_L$. Then we know that $l'_L \cdot o \in \textit{PQueue}$. Since $l'_R$ is obtained from $l'$ by discarding pairs of matched $\textit{put}$ and $\textit{rm}$ operations, it is easy to see that $L'_R \in \textit{PQueue}$, and then we know that $l'' = l'_L \cdot o \cdot L'_R \in \textit{PQueue}$.

It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \textit{PQueue}$. \qed
\end {proof}

The following lemma states that $\textit{PQueue}$ is step-by-step linearizability, it is a direct consequence of Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, Lemma \ref{lemma:PQ2Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ2Equal is step-by-step linearizability} and Lemma \ref{lemma:PQ3 is step-by-step linearizability}.

@@

\begin {proof}
This is a direct consequence of Lemma \ref{lemma:PQ1Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ1Equal is step-by-step linearizability}, Lemma \ref{lemma:PQ2Lar is step-by-step linearizability}, Lemma \ref{lemma:PQ2Equal is step-by-step linearizability} and Lemma \ref{lemma:PQ3 is step-by-step linearizability}. \qed
\end {proof}

\subsection{Proof of Lemma \ref{lemma:PQ as multi in MRpri for history}}

\PQasMultiInMRpriforHistory*

To prove the $\textit{only if}$ direction, assume that $e \sqsubseteq l \in \textit{PQueue}$. Given $e' = e \vert_{D}$, by Lemma \ref{lemma:PQ is closed under projection}, it is obvious that $e' \sqsubseteq l' = l \vert_{D} \in \textit{PQueue}$. Then by Lemma \ref{lemma:PQ as multi in MRpri for sequence} we know that $l' \in \textit{MS}(R)$ where $R = \textit{last}(e')$.

To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{last}(e) = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{=}$, and the maximal priority of $e$ is unmatched $\textit{put}$ with items in $S$ and matched $\textit{put}$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{=}$, and the maximal priority in $e$ is of more than one pair of matched $\textit{put}$ and $\textit{rm}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$, where such $\textit{put}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$, and such $\textit{rm}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{rm}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{>}$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$ with maximal priority in $e$.
\end{itemize}

Similarly, we obtain $e_2$ from $e_1$, obtain $e_3$ from $e_2$, $\ldots$, until we obtain some $e_m = \epsilon$. Let $e_0=e$. It is obvious that $e_m \sqsubseteq \textit{PQueue}$. For $e_{\textit{m-1}}$, since

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $e_m \sqsubseteq \textit{PQueue}$,

\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} \in \{ \textit{PQ}_1^{>}, \textit{PQ}_1^{=}, \textit{PQ}_2^{>}, \textit{PQ}_2^{=} \}$ and $e$ matches $R_{\textit{m-1}}$ with witness $x$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_{\textit{m-1}} \setminus x \sqsubseteq \textit{PQueue}$, and by step-by-step linearizability of $\textit{PQueue}$, we can see that $e_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.

\item[-] If $\textit{last}(e_{\textit{m-1}}) = R_{\textit{m-1}} = \textit{PQ}_3$ and $o$ is a $\textit{rm}(\textit{empty})$ in $e_{\textit{m-1}}$: We already know that $e_{\textit{m-1}} \sqsubseteq \textit{MS}(R_{\textit{m-1}})$, $e_{\textit{m-1}} \setminus o \sqsubseteq \textit{PQueue}$, and by step-by-step linearizability of $\textit{PQueue}$, we can see that $e_{\textit{m-1}} \sqsubseteq \textit{PQueue}$.
\end{itemize}

@@
